---
title: "RegionalRetina_Figures"
author: "Drew Voigt"
date: "3/17/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

Please see the RegionalRetina_aggregating_filtering_classifying.Rmd script or the hosted seurat RData object at https://drive.google.com/drive/folders/1HWB7T4PqWSyYeBX3uVdAuRCmzYqeDXgA?usp=sharing for access the the seurat object used for these scripts.
```{r}
library(Seurat) # v3.2.3
library(tidyverse)
load("data/data_from_google_drive/seurat_object.RData")
```

Figure 3A: dimensionality reduction plot
```{r}
gg_color_hue <- function(n, light_intensity) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = light_intensity, c = 100)[1:n]
}

light <- gg_color_hue(15,87)
dark <- gg_color_hue(15,65)

class_colors <- c(dark[13], light[12],
                  dark[11], light[11],
                  dark[9], light[8],
                  dark[7], light[5],
                  dark[4], light[3],
                  dark[2], light[2],
                  dark[1], light[1]
                  )
Idents(seurat_obj) <- "celltype"
DimPlot(seurat_obj, cols = (class_colors))
```


Figure 3B: Comparison of cell proportions
```{r}
celltype_levels <- seurat_obj@meta.data %>% 
  .[["celltype"]]

seurat_obj[["celltype_levels"]] <- factor(celltype_levels, 
                                          levels = c("rod", "cone", "cone-off-BC",
                                                     "cone-off-BC-BC3A", "cone-on-BC", "rod-BC",
                                                     "RGC", "amacrine", "horizontal", "Muller", "astrocyte", 
                                                     "microglia", "pericyte", "endothelial"))


proportion_data <- seurat_obj@meta.data %>% 
  select(library, celltype_levels) %>%
  group_by(library, celltype_levels) %>%
  summarize(n = n()) %>%
  mutate(freq = n / sum(n)) %>%
  ungroup() %>%
  complete(library, celltype_levels, fill = list(n = 0, freq = 0)) %>%
  mutate(region = ifelse(str_detect(library, "perifovea"), "perifovea", "fovea"))


proportion_data %>% 
    ggplot(aes(x = celltype_levels, 
               y = freq, 
               fill = celltype_levels, 
               group = library)) +
      geom_bar(
        stat = "identity",
        position = "dodge",
        color = "black"
      ) +
    scale_fill_manual(values = (class_colors)) + 
    coord_flip() + 
    facet_wrap(~region, nrow = 1) +
  theme(legend.position = "none",
        axis.title = element_text(size = 12),
        axis.text = element_text(size = 12))
```


Figure 3C: Violin Plots
```{r}
## using violin plotting functions from our package cellcuratoR (github.com/cellcuratoR)
source(here("scripts/github_plotting_functions.R"))
overview_vln_genes <- c("PDE6A", "ARR3", "GRIK1", "ERBB4", "GRM6", 
                        "PRKCA", "PRPH", "GAD1", "ONECUT2", "RLBP1", 
                        "GFAP", "C1QA", "RGS5", "VWF")

seurat_obj[["final_cluster_labels"]] <- seurat_obj[["celltype"]] ## this function requires an input column called 'final_cluster_labels'
construct_violin_plot(my_object = seurat_obj,
                                  genes_to_investigate = overview_vln_genes,
                                  colors = class_colors)
```

Figure 4: Regional expression in CONE PHOTORECEPTOR CELLS
```{r}
load("data/data_from_google_drive/retina_by_region_6_studies.RData")
# retina_by_region is a list with six elements
# Each list element contains differential expression results from six independent scRNA-seq
# studies of human retina with at least two regional libraries. 
# element 1: current study
# element 2: Voigt et al foveal vs peripheral retina, 2019
# element 3: Yan et al foveal vs peripheral retina, 2020
# element 4: Cowan et al foveal vs peripheral retina, 2020
# element 5 Liang et al macular vs peripheral retina, 2020 (single-nucleus RNA-seq)
# element 6: Sridhar et al central vs peripheral INFANT retina, 2020

# Details on processing of each complementary study can be provided upon request. 

source(here("scripts/github_plotting_functions.R"))

my_factor_levels <- c("bullseye_fov_perifov",
                           "EER_fov_periph",
                           "yan_fov_periph",
                           "cowan_fov_per", 
                           "liang_mac_periph",
                           "sridhar_cent_periph")
my_colors <- c("#d73027",
"#fc8d59",
"#fee090",
"#e0f3f8",
"#91bfdb",
"#4575b4")

cone_genes <- c("RIMS2", "CPLX4", "WRB", "SLC4A7", "POU2AF1", "GPATCH1", "COL4A3", "BCO2", "CC2D2A", "RPGR", "RP1L1", "BICD1", "WDR47")
for(i in 1:length(cone_genes)){
  regional_figure(object = seurat_obj,
                  cell_population = "cone",
                  celltype_vector = rep("cone", 6),
                  logFC_variable = "logFC.pseudobulk",
                  gene_of_interest = cone_genes[i],
                  dge_list = retina_by_region,
                  dataset_factor_levels = my_factor_levels,
                  dataset_colors = my_colors,
                  upper_ylim = 3)

}

# if the logFC is so high that it is off scale, you will see a black bar with the logFC numerically shown below the bar
```


Figure 5: Transcription factors enriched in foveal cones
```{r}
## TFs identified from the SysGo database

# foveal enriched TFs
cone_tfs <- c("POU2AF1", "YBX1", "LBH", "FLYWCH2", "ZNF358", "ZBTB38",
              "ZC3H6", "THAP5", "ZCRB1", "RCAN3", "BHLHE41", "ZC2HC1A",
              "YAF2", "ZC3H15", "CEBPZ", "MIER1", "ZNF385B", "VEZF1",
              "PHF19", "GTF3A")
for(i in 1:length(cone_tfs)){
  regional_figure(object = seurat_obj,
                  cell_population = "cone",
                  celltype_vector = rep("cone", 6),
                  logFC_variable = "logFC.pseudobulk",
                  gene_of_interest = cone_tfs[i],
                  dge_list = retina_by_region,
                  dataset_factor_levels = my_factor_levels,
                  dataset_colors = my_colors,
                  upper_ylim = 3)

}
# if the logFC is so high that it is off scale, you will see a black bar with the logFC numerically shown below the bar

```


SI Figure 3: Transcription factors enriched in peripheral cones
```{r}
# perifoveal enriched TFs
cone_tfs <- c("IER5", "TEAD3", "ZNRF1", "MBD5", "BAZ2B", "AFF1", "ZNF207", "REL", "NFKBIA",
              "TBPL1", "ADNP2", "ZNF274", "ZNF451", "GPBP1", "RAX", "GTF3C1", "FOSB", "ZC3H14", "KLF6", "ZFYVE28")
for(i in 1:length(cone_tfs)){
  regional_figure(object = seurat_obj,
                  cell_population = "cone",
                  celltype_vector = rep("cone", 6),
                  logFC_variable = "logFC.pseudobulk",
                  gene_of_interest = cone_tfs[i],
                  dge_list = retina_by_region,
                  dataset_factor_levels = my_factor_levels,
                  dataset_colors = my_colors,
                  upper_ylim = 3)

}
# if the logFC is so high that it is off scale, you will see a black bar with the logFC numerically shown below the bar
```


Figure 6: Regional expression in ROD PHOTORECEPTOR CELLS and MULLER CELLS
```{r}
load("data/data_from_google_drive/retina_by_region_6_studies.RData")
# retina_by_region is a list with six elements
# Each list element contains differential expression results from six independent scRNA-seq
# studies of human retina with at least two regional libraries. 
# element 1: current study
# element 2: Voigt et al foveal vs peripheral retina, 2019
# element 3: Yan et al foveal vs peripheral retina, 2020
# element 4: Cowan et al foveal vs peripheral retina, 2020
# element 5 Liang et al macular vs peripheral retina, 2020 (single-nucleus RNA-seq)
# element 6: Sridhar et al central vs peripheral INFANT retina, 2020

source(here("scripts/github_plotting_functions.R"))

my_factor_levels <- c("bullseye_fov_perifov",
                           "EER_fov_periph",
                           "yan_fov_periph",
                           "cowan_fov_per", 
                           "liang_mac_periph",
                           "sridhar_cent_periph")
my_colors <- c("#d73027",
"#fc8d59",
"#fee090",
"#e0f3f8",
"#91bfdb",
"#4575b4")

rod_genes <- c("PNO1", "RBP7", "SCAPER", "AHI1")
for(i in 1:length(rod_genes)){
  regional_figure(object = seurat_obj,
                  cell_population = "rod",
                  celltype_vector = rep("rod", 6),
                  logFC_variable = "logFC.pseudobulk",
                  gene_of_interest = rod_genes[i],
                  dge_list = retina_by_region,
                  dataset_factor_levels = my_factor_levels,
                  dataset_colors = my_colors,
                  upper_ylim = 3)

}

muller_genes <- c("FAT1", "DAPL1", "CYP26A1", "TF", "MXRA7", "RDH10", "RLBP1", "FABP5")
for(i in 1:length(muller_genes)){
  regional_figure(object = seurat_obj,
                  cell_population = "Muller",
                  celltype_vector = c(rep("Muller", 4), "muller", "Muller"),
                  logFC_variable = "logFC.pseudobulk",
                  gene_of_interest = muller_genes[i],
                  dge_list = retina_by_region,
                  dataset_factor_levels = my_factor_levels,
                  dataset_colors = my_colors,
                  upper_ylim = 3)

}

# if the logFC is so high that it is off scale, you will see a black bar with the logFC numerically shown below the bar
```




