---
title: "RegionalRetina_aggregating_filtering_classifying"
author: "Voigt"
date: "3/17/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---


This Rmd script will overview how to aggregate count files into a Seurat object. Data will be filtered in an identical manner to the manuscript.

In addition to the raw count files, you can also download a pre-aggregated Seurat object .RData file from XXXX and skip this script.

```{r}
library(Seurat) # v3.2.3
library(tidyverse)
```

scRNA-seq data can be downloaded from GEO with the accession number GSE169047.
Raw count files have the extension "_counts.csv"
```{r}
my_files <- list.files("~/LSS/IVR/NGS/raw_data/20200821_uiowa_bullseye_retina/geo/")
my_count_files <- my_files[grepl(pattern = "*counts.csv", x = my_files)]

# create a list to hold seurat objects from each library
my_object_list <- vector(length = length(my_count_files), mode = "list")

for(i in 1:length(my_count_files)){
  count_matrix <- read_csv(paste0("~/LSS/IVR/NGS/raw_data/20200821_uiowa_bullseye_retina/geo/", my_count_files[i]))
  
  meta_data <- count_matrix[,c("barcode", "region", "celltype", "library")]
  
  # peel off metadata for creation of 10X object
  # this matrix must be transposed (genes x barcodes) to work with CreateSeuratObject
  count_matrix <- count_matrix[, -c(1:4)] %>%
    as.matrix() %>%
    t()
  colnames(count_matrix) <- meta_data[["barcode"]]
  
  # pull of the donor number and region for this donor
  donor_information <- meta_data[["library"]][1]
  
  my_initial_object <- CreateSeuratObject(counts = count_matrix,
                                          project = donor_information,
                                          assays = "RNA")
  
  # identify the percent of mitochondrial genes
  my_initial_object[["mito.genes"]] <- PercentageFeatureSet(my_initial_object, pattern = "^MT-") #
  
  my_object_list[[i]] <- my_initial_object
  names(my_object_list)[[i]] <- donor_information
}
```

Next, we will filter the object for lowly expressed genes (low quality cells), highly expressed genes (doublets), or cells with a high percent of mitochondrial genes (low quality cells). We will use the filtering criteria from the methods sections of the manuscript.
```{r}
lower.cutoff = 400
higher.cutoff = 6500
mito.cutoff = 60
selection_method = "vst"
variable_genes = 2000

for(i in 1:length(my_object_list)){
  my_initial_object <- my_object_list[[i]]
  my_intial_object <- subset(my_initial_object, 
                             subset = 
                               nFeature_RNA < higher.cutoff & 
                               nFeature_RNA > lower.cutoff & 
                               mito.genes < mito.cutoff
                             )
  my_intial_object <- NormalizeData(my_intial_object, normalization.method = "LogNormalize", scale.factor = 10000)
  my_intial_object <- FindVariableFeatures(my_intial_object, selection.method = selection_method, nfeatures = variable_genes)
  my_object_list[[i]] <- my_intial_object
}
```


Next, we perform cannonical correlation analysis and find anchors.
This is computationally demanding.
```{r}
object.anchors <- FindIntegrationAnchors(object.list = my_object_list, dims=1:25)
object.combined <- IntegrateData(anchorset = object.anchors, dims = 1:25)

DefaultAssay(object.combined) <- "integrated"
object.combined <- ScaleData(object.combined, verbose = FALSE)
object.combined <- RunPCA(object.combined, npcs = 30, verbose = FALSE)
```

Finally, we perform the clustering and dimensionality rediction
```{r}
ElbowPlot(object.combined, ndims = 25)
final.dim = 23 # 23 seems like a reasonable number of dimensions
  
object.combined <- RunUMAP(object.combined, reduction = "pca", dims = 1:final.dim)
object.combined <- FindNeighbors(object.combined, reduction = "pca", dims = 1:final.dim)
object.combined <- FindClusters(object.combined, resolution = 0.5)
```

Finally, we classify clusters into cells based on expression of known marker genes.
```{r}
DefaultAssay(object.combined) <- "integrated"
object.combined <- FindClusters(object.combined, resolution = 0.4)
DimPlot(object.combined, label = TRUE)

DefaultAssay(object.combined) <- "RNA"
FeaturePlot(object.combined, features = "RHO") # rod
FeaturePlot(object.combined, features = "ARR3") # cone
FeaturePlot(object.combined, features = "GRM6") # CONE-on-BC
FeaturePlot(object.combined, features = "PRKCA") # ROD-BC
FeaturePlot(object.combined, features = "GRIK1") # cone-off-BC
FeaturePlot(object.combined, features = "ONECUT2") # HZ
FeaturePlot(object.combined, features = "GAD1") # Amacrine
FeaturePlot(object.combined, features = "C1QA") # microglia
FeaturePlot(object.combined, features = "VWF") # endothelial
FeaturePlot(object.combined, features = "ACTA2") # pericyte
FeaturePlot(object.combined, features = "RLBP1") # Muller
FeaturePlot(object.combined, features = "GFAP") # Astrocyte
FeaturePlot(object.combined, features = "THY1") # RGC



DefaultAssay(object.combined) <- "RNA"
ordered_celltypes <- c("muller", "rod", "muller", "amacrine", "rgc", 
                       "rod-bipolar", "cone-off-bipolar", "cone-on-bipolar", "cone-on-bipolar", "cone-off-bipolar", 
                       "cone-on-bipolar", "horizontal", "cone-off-bipolar", "rgc", "endothelial",
                       "cone", "cone-off-bipolar", "astrocyte", "rgc", "pericyte", 
                       "muller", "microglia", "rgc")

celltype <- plyr::mapvalues(object.combined[["seurat_clusters"]][["seurat_clusters"]],
                            from = seq(from = 0, to = 22),
                            to = ordered_celltypes)

object.combined[["celltype"]] <- celltype
```



Save your object
```{r}
# save(object.combined, file = "save_me_somewhere.RData")
```



